---
title: 'EPID700 GitHub Practicum'
author: "Reproducible Research with R and GitHub"
output: html_document
---
***

#### Load and Describe Data
We will use an NHANES file . Refer to its [data dictionary](https://github.com/HimesGroup/EPID600_2016/blob/master/DataFiles/NHANES_2007to2008_DataDictionary.md) to understand its contents. We will use the [`dplyr()`](https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html) package, a relatively new package by Hadley Wickham of RStudio that facilitates managing of data frames, to modify the file after loading. Install `dplyr` using `install.packages("dplyr")` if you need to. 

```{r, eval=TRUE, message=FALSE}
library(dplyr)
nhanes <- read.csv(url("https://raw.githubusercontent.com/HimesGroup/EPID600_2016/master/DataFiles/NHANES_2007to2008.csv"), header=TRUE)
nhanes <- rename(nhanes, id=SEQN, gender=RIAGENDR, age=RIDAGEYR, race=RIDRETH1, education=DMDEDUC2, income=INDHHIN2, health.provider=HUQ040, wheezing=RDQ070, asthma=MCQ010, voc=WTSVOC2Y, fvc=SPXNFVC, fev1=SPXNFEV1) #Renaming columns
str(nhanes)
nhanes <- mutate(nhanes, asthma=ifelse(asthma %in% c(7,9), NA, ifelse(asthma==2, 0, 1))) #Changing definition of asthma so that sums of variable are equal to number of cases by setting cases==1, controls==0, else NA
nhanes <- mutate(nhanes, gender = factor(gender, levels=c(1, 2), labels=c("male", "female")))
income.dis <- nhanes %>%
                filter(income %in% c(1:10, 14, 15)) %>%
                mutate(income.short = cut(income, breaks=c(0, 5, 10, 15), labels=c("<$25K", "$25-75K", ">$75K"))) %>%
                count(gender, income.short)
income.dis
head(nhanes)
```

To get a sense of the distribution of a categorical variable, barcharts are helpful.
```{r, eval=TRUE}
barplot(table(nhanes$gender))
barplot(table(nhanes$race))
```

To get a sense of the distribution of a continuous variable, histograms are helpful.
```{r, eval=TRUE}
hist(nhanes$age)
```

#### Create Data File

Here we will generate random data for a few probability distributions. For normally distributed variables, we can use `rnorm`, which generates variates according to the Normal distribution given a mean and standard deviation. For binary outcome variables, we can use `rbinom`, which generates variates according to the binomial distribution given _n_ independent experiments with success probability _p_. We will use `set.seed()` to ensure that the same numbers are generated for a call.

```{r, eval=TRUE}
set.seed(1234)
x.normal <- rnorm(1000)
e <- rnorm(1000, 0, 2)
y.normal <- 0.5 + 2*x.normal + e
summary(y.normal)
plot(x.normal, y.normal)
```

If x were a binary outcome variable, like gender, we could simulate it using `rbinom`, which corresponds to the binomial distribution.
```{r, eval=TRUE}
x.binomial <- rbinom(1000, 1, 0.5)
y.binomial <- 10 + 3*x.binomial + e
summary(y.binomial)
boxplot(y.binomial ~ x.binomial)
```

#### Analyze and Plot Data

