---
title: 'EPID700 GitHub Practicum'
author: "Reproducible Research with R and GitHub"
output: html_document
---
***

##### testing 

This file has R commands that are often used to analyze epidemiology data. They are by no means comprehensive, but if you are new to R, they are a good set to get started. Run parts of them individually on your own computer, and add code where requested. When you are done with your edits, use the `Knit HTML` button above to make an html version of this file. When you are done, select the `Git` tab in the right upper panel, use the `pull` command to make sure any remote changes are saved locally, then `stage` the changed files you would like to move to GitHub, select `Commit` with a message about these changes, and then `push` the changes to your GitHub repository. To finish the assignment, remember to make a pull request on the original EPID700 repository from your own copy on GitHub.


### Load and Describe Data
We will use an abbreviated [NHANES](http://www.cdc.gov/nchs/nhanes/) file that is saved in a [GitHub repository](https://raw.githubusercontent.com/HimesGroup/EPID600_2016/master/DataFiles/NHANES_2007to2008.csv). You may want to refer to its [data dictionary](https://github.com/HimesGroup/EPID600_2016/blob/master/DataFiles/NHANES_2007to2008_DataDictionary.md) to understand its contents. We will use the [`dplyr()`](https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html) package, a relatively new package by Hadley Wickham of RStudio that facilitates managing of data frames, to modify the file after loading. Install `dplyr` using `install.packages("dplyr")` if you need to. 


```{r, eval=TRUE, message=FALSE}
library(dplyr)
nhanes <- read.csv(url("https://raw.githubusercontent.com/HimesGroup/EPID600_2016/master/DataFiles/NHANES_2007to2008.csv"), header=TRUE)
nhanes <- rename(nhanes, id=SEQN, gender=RIAGENDR, age=RIDAGEYR, race=RIDRETH1, education=DMDEDUC2, income=INDHHIN2, health.provider=HUQ040, wheezing=RDQ070, asthma=MCQ010, voc=WTSVOC2Y, fvc=SPXNFVC, fev1=SPXNFEV1) #Renaming columns
nhanes <- mutate(nhanes, asthma = factor(asthma, levels=c(1, 2), labels=c("asthma", "non-asthma"))) 
nhanes <- mutate(nhanes, gender = factor(gender, levels=c(1, 2), labels=c("male", "female")))
str(nhanes)
```

Plots are helpful as part of exploratory analyses to get a sense of the distribution of variables. Here are plots of gender and age distributions of this dataset:
```{r, eval=TRUE}
barplot(table(nhanes$gender))
hist(nhanes$age)
```

> Insert a code chunk here to (1) make a table of gender counts by asthma status and (2) create a barplot of how gender is distributed by asthma status


### Simple Regression

Here we will generate random data for a few probability distributions. For normally distributed variables, we can use `rnorm`, which generates variates according to the Normal distribution given a mean and standard deviation. For binary outcome variables, we can use `rbinom`, which generates variates according to the binomial distribution given _n_ independent experiments with success probability _p_. We will use `set.seed()` to ensure that the same numbers are generated for a call.

```{r, eval=TRUE}
set.seed(1234)
x.normal <- rnorm(1000) #A normally distributed variable
e <- rnorm(1000, 0, 2) #Give fluctuations to a distribution with a fake error term
y.normal <- 0.5 + 2*x.normal + e #A linear outcome variable that is dependent on x.normal
summary(x.normal)
summary(y.normal)
plot(x.normal, y.normal)

x.binomial <- rbinom(1000, 1, 0.5) #A binomially distributed variable
y.binomial <- 10 + 3*x.binomial + e #An outcome variable that is dependent on x.normal
summary(x.binomial)
summary(y.binomial)
boxplot(y.binomial ~ x.binomial)
```

> Insert a code chunk here to model the relationship between (1) x.normal and y.normal and (2) x.binomial and y.binomial using the `lm` function, and display the output using `summary`. Use `?lm` if you are not familiar with this function.

